---
title: "Sunfish_map"
author: "Johan Cronje"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

loading the required packages
```{r}
library(leaflet)
library(htmltools)
library(RColorBrewer)
library(rgbif)
library(tidyverse)
library(sf)
```


Importing a subset of the data, containing just the fish family I am interested in, for the map.
```{r}
#download key
res <- "0003331-260225131425191"

centrarchidae<- occ_download_get(res) %>% occ_download_import() %>% 
  filter(family == 'Centrarchidae')
```

Downloaded dataset has coordinates but isn't seen as a spatial object by R, just as a dataset. This code converts it into a spatial object needed to plot it on a map
```{r}
centrarchidae<- st_as_sf(centrarchidae, coords = c('decimalLongitude','decimalLatitude'), crs =4326)

```

creating a color palette for the different species that will be plotted on the map
```{r}
pal <- c('#8F00FF','#FB00FF','#00D2FF','#00FF47')
```

This piece of code creates an interactive map and plots where each species of centrachidae found in South africa have occured.
```{r}
leaflet() %>%
  addTiles(group = 'specificEpithet') %>%
  addCircleMarkers(data =  centrarchidae,
                   group =c("macrochirus","dolomieu",'punctulatus','salmoides'),
                   radius = 0.5,
                   color = pal,
                   opacity =1) %>%
  addLegend(position = "topleft",
            color=pal,
            labels = c("Bluegill","Smallmouth Bass",'Spotted Bass','Largemouth Bass (Florida var.)')) 
```

Figure 1. Map showing the occurence of four species of fish in the family Centrarchidae across South Africa.  

